---
title: "Filtering Pipeline"
author: "Christoph Petrynowski"
date: "2023-02-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Software-Library

```{r library, echo=FALSE}
library(PERFect)
library(phyloseq)
library(decontam)
library(tidyverse)
library(pairwiseAdonis)
library(kableExtra)
library(vegan)
library(tidytree)
library(ape)
library(ggrepel)
library(microbiome)
library(ggpubr) # for manual p-value assigning
library(rstatix) # for wilcox test
library(yingtools2)
#if( !require(ggtree) ){
#  install.packages(ggtree)
#}
library(readxl) # load excel data
if( !require(rlist) ){
  install.packages(rlist)
}
library(rlist) # work with lists
if( !require(ggsignif) ){
  install.packages(ggsignif)
}
library(ggsignif) # plot p.values on a ggplot
set.seed(3)
library(plyr) #for ldply - performs function on every element in list & combines the output into a dataframe
library(reshape2) # for melting data
if( !require(doParallel) ){
  install.packages("doParallel")
}
library(doParallel) # ConQuR´s GitHub recomends this step
#devtools::install_github("wdl2459/ConQuR")
library(ConQuR)
if( !require("VennDiagram") ){
  install.packages("VennDiagram")
}
library("VennDiagram")
if( !require("ggvenn") ){
  install.packages("ggvenn")
}
library("ggvenn")
library(RColorBrewer) 
```

```{r}
## adding functions
source("../CP/MyFunctionsV2.R")
source("../CP/MyAnalysisWithPlots.R")
```

# Raw Phyloseq 1500.80s
### Loading phyloseq "1500.80s true samples" objects
```{r loading phyloseq "1500.80s true samples" objects}
physeq1.1500.80 <- readRDS("../raw_data/PhyloSeqObj/2022_05_27_patho_FFPE_tum_panc_1/physeq_original_1500.80.rds")
physeq2.1500.80 <- readRDS("../raw_data/PhyloSeqObj/2022_05_27_patho_FFPE_tum_panc_2/physeq_original_1500.80.rds")
physeq3.1500.80 <- readRDS("../raw_data/PhyloSeqObj/2022_05_28_patho_FFPE_tum_panc_3/physeq_original_1500.80.rds")
physeq4.1500.80 <- readRDS("../raw_data/PhyloSeqObj/2022_05_28_patho_FFPE_tum_panc_4/physeq_original_1500.80.rds")
physeq5.1500.80 <- readRDS("../raw_data/PhyloSeqObj/2021_12_01_patho_FFPE_self_tumor/physeq_original_1500.80.rds")

#physeq6.1500.80 <- readRDS("raw_data/PhyloSeqObj/2022_08_09_patho_FFPE_tum_panc_rep1/physeq_original_1500.80.rds") #09.08.22: rep1: 10 samples (tumor+normal panc mixed)
#physeq7.1500.80 <- readRDS("raw_data/PhyloSeqObj/2022_08_09_patho_FFPE_tum_panc_rep2/physeq_original_1500.80.rds") #09.08.22: rep2: 10 samples (tumor+normal panc mixed)
p.true.original.1500.80 <- merge_phyloseq(physeq1.1500.80,physeq2.1500.80,physeq3.1500.80,physeq4.1500.80,physeq5.1500.80)#merge the different phyloseq objects into 1 phyloseq object
# sample_data of new phyloseq-object

p.true.filt.1500.80 <- subset_taxa(p.true.original.1500.80, superkingdom=="Bacteria")
p.true.filt.s.1500.80 <- tax_glom(p.true.filt.1500.80, taxrank = "species", NArm = TRUE)
rm(list = c("physeq1.1500.80", "physeq2.1500.80", "physeq3.1500.80", "physeq4.1500.80", "physeq5.1500.80"))
rm(p.true.filt.1500.80)
```

### Loading phyloseq "1500.80s ntc samples" objects
```{r loading phyloseq "1500.80s ntc samples" objects}
#loading 1500.80s ntc samples
physeq.nct1.1500.80 <- readRDS("../raw_data/PhyloSeqObj/2022_06_02_patho_FFPE_tum_panc_ntc_1/physeq_original_1500.80.rds")
physeq.nct2.1500.80 <- readRDS("../raw_data/PhyloSeqObj/2022_06_02_patho_FFPE_tum_panc_ntc_2/physeq_original_1500.80.rds")
physeq.nct3.1500.80 <- readRDS("../raw_data/PhyloSeqObj/2022_06_03_patho_FFPE_tum_panc_ntc_3/physeq_original_1500.80.rds")
physeq.nct4.1500.80 <- readRDS("../raw_data/PhyloSeqObj/2022_06_03_patho_FFPE_tum_panc_ntc_4/physeq_original_1500.80.rds")
physeq.nct5.1500.80 <- readRDS("../raw_data/PhyloSeqObj/2022_06_04_patho_FFPE_tum_panc_ntc_5/physeq_original_1500.80.rds")
physeq.nct6.1500.80 <- readRDS("../raw_data/PhyloSeqObj/2022_06_04_patho_FFPE_tum_panc_ntc_6/physeq_original_1500.80.rds")
physeq.nct7.1500.80 <- readRDS("../raw_data/PhyloSeqObj/2022_06_10_patho_FFPE_tum_panc_ntc_7/physeq_original_1500.80.rds")
physeq.nct8.1500.80 <- readRDS("../raw_data/PhyloSeqObj/2022_07_13_patho_FFPE_tum_panc_ntc_8/physeq_original_1500.80.rds")
physeq.nct9.1500.80 <- readRDS("../raw_data/PhyloSeqObj/2022_07_13_patho_FFPE_tum_panc_ntc_9/physeq_original_1500.80.rds")
physeq.nct10.1500.80 <- readRDS("../raw_data/PhyloSeqObj/2021_12_02_patho_FFPE_self_ntc/physeq_original_1500.80.rds")
physeq.nct11.1500.80 <- readRDS("../raw_data/PhyloSeqObj/2022_01_24_patho_FFPE_self_ntc_2/physeq_original_1500.80.rds")

physeq.nct.original.1500.80 <- merge_phyloseq(physeq.nct1.1500.80, physeq.nct2.1500.80, physeq.nct3.1500.80, physeq.nct4.1500.80, physeq.nct5.1500.80, physeq.nct6.1500.80, physeq.nct7.1500.80, physeq.nct8.1500.80, physeq.nct9.1500.80,physeq.nct10.1500.80, physeq.nct11.1500.80)

physeq.filt.nct.1500.80 <- subset_taxa(physeq.nct.original.1500.80, superkingdom=="Bacteria")
physeq.filt.nct.s.1500.80 <- tax_glom(physeq.filt.nct.1500.80, taxrank = "species", NArm = TRUE)

rm(list = c('physeq.nct1.1500.80', 'physeq.nct2.1500.80', 'physeq.nct3.1500.80', 'physeq.nct4.1500.80', 'physeq.nct5.1500.80',
            'physeq.nct6.1500.80', 'physeq.nct7.1500.80', 'physeq.nct8.1500.80', 'physeq.nct9.1500.80', 'physeq.nct10.1500.80', 'physeq.nct11.1500.80'))
rm(physeq.filt.nct.1500.80)

#replace sample_side "paraffin" with "paraffin_tumor"
sample_data(physeq.filt.nct.s.1500.80)$sample_side <- ifelse(sample_data(physeq.filt.nct.s.1500.80)$sample_side == "paraffin", "paraffin_tumor", sample_data(physeq.filt.nct.s.1500.80)$sample_side)
```

Before using NCTs for decontamination, we firstly need to remove samples from NCTs whose read counts are low.

Overview of read counts in samples in NCTs

```{r}
visualize_samples_low_reads(physeq.filt.nct.s.1500.80)
```

We decide to remove NCT samples if their read counts are less than 150

```{r}
#physeq.filt.nct.s.1500.80 <- filter_samples_low_reads(PhyObj = physeq.filt.nct.s.1500.80,threshold = 150)
bac.count.ntc <- tibble(bac.count = sample_sums(physeq.filt.nct.s.1500.80), 
                          sample = sample_data(physeq.filt.nct.s.1500.80)$id)
id.ex <- bac.count.ntc %>% filter(bac.count < 150) %>% pull (sample)
physeq.read.filt <- subset_samples(physeq.filt.nct.s.1500.80, !(id %in% id.ex))
physeq.read.filt <- prune_taxa(taxa_sums(physeq.read.filt)>0, physeq.read.filt)
physeq.filt.nct.s.1500.80 <- physeq.read.filt
rm(bac.count.ntc)
rm(id.ex)
rm(physeq.read.filt)
```

Now, in NTC samples, we have *`r nrow(sample_data(physeq.filt.nct.s.1500.80))`* samples and *`r nrow(otu_table(physeq.filt.nct.s.1500.80))`* taxa.

#OTU unification

We have 2 phyloseq objects that may describe the same species using different OTU handles. We therefore unify the OTU handles for an easier downstream analysis.

### Merge TRUE/NCT samples
```{r merge true/nct samples}
p.comb.filt.s.1500.80 <- merge_phyloseq(p.true.filt.s.1500.80,physeq.filt.nct.s.1500.80) # merge the phyloseq objects containing true samples (p.true.filt.s.1500.80) and nct samples (physeq.filt.nct.s.1500.80)
p.comb.filt.s.1500.80 <- tax_glom(p.comb.filt.s.1500.80,taxrank = "species", NArm = TRUE) # agglomerate taxa to species level. This combines different OTUs {hypothecially: 450 / 453} assigned to the same species into just 1 OTU
sample_data(p.comb.filt.s.1500.80)$is.neg <- sample_data(p.comb.filt.s.1500.80)$TRUE_control == "control"#first identify your negative controls -> decontam!

p.comb.filt.s.1500.80 <- subset_samples(p.comb.filt.s.1500.80, uniqueID != 326 & uniqueID != 327 & uniqueID != 328 & uniqueID != 329 & uniqueID != 330 & uniqueID != 331 & uniqueID !=332 & uniqueID !=333 & uniqueID !=334 & uniqueID !=335 & uniqueID !=338 & uniqueID !=339 & uniqueID !=340 & uniqueID !=341)
```


## DEMERGE!
```{r}
p.true.filt.s.1500.80.unify <- subset_samples(p.comb.filt.s.1500.80,TRUE_control=="TRUE") #only true samples
p.true.filt.s.1500.80.unify <- prune_taxa(x = p.true.filt.s.1500.80.unify,taxa = taxa_names(p.true.filt.s.1500.80.unify)[-(which(taxa_sums(p.true.filt.s.1500.80.unify)==0))]) #only taxa found in true samples

physeq.filt.nct.s.1500.80.unify <- subset_samples(p.comb.filt.s.1500.80,TRUE_control == "control") #only NCT samples
physeq.filt.nct.s.1500.80.unify <- prune_taxa(x = physeq.filt.nct.s.1500.80.unify,taxa = taxa_names(physeq.filt.nct.s.1500.80.unify)[-(which(taxa_sums(physeq.filt.nct.s.1500.80.unify)==0))]) #only taxa found in NCT samples
```

### Low abundance filtering (Step 3)

In this step, we remove all taxas in true samples whose abundance are low (eg. \<10\^-4) in atleast 1 sample
```{r}
p.true.RmLowAbun.man <- filter_by_low_abundance(phyloseqObj = p.true.filt.s.1500.80.unify,threshold =0.0001 ,A = 1)
```

First of all, let us check the taxa overlap between NCT samples and true sample after low-abundance filtering by rule of thumb.

```{r}
PlotVenn2Sets(taxa_names(p.true.RmLowAbun.man) %>% unique(),
              taxa_names(physeq.filt.nct.s.1500.80.unify) %>% unique(),
              "True Samples" , "NCT Samples ")
```

There are `r setdiff(taxa_names(physeq.filt.nct.s.1500.80.unify), taxa_names(p.true.RmLowAbun.man)) %>% length()` taxa belong to NCT samples but not in true samples (after rule of thumb). They are obvious contamination.

Instead of Decontam R package, we follow procedure of Nejman et al..

# Nejman

#### High prevalence taxas in NCTs

We consider all taxas whose prevalence in NCTs are higher than a certain threshold (eg. 30%) are contaminants.

In order to determine that threshold, we have a look at the plot below

```{r}
xout <- HighPrevalence_Data(p.true.RmLowAbun.man, physeq.filt.nct.s.1500.80.unify)
plot(xout$factorx, xout$rel_tax, xlab="prevalence in NCT samples", ylab="proportion of tax in true samples")
```

This plot shows us how many percents of taxa we would remove from true samples (y-axis) if we pick up a threshold for high prevalence in NCT samples (x-axis).

From this plot, a threshold 0.3 would be fine.
Taking a filter of 0.307 yields a loss of 12 taxa or roughly 1.94%

```{r}
filter <- 0.339
filtered_taxa1 <- tibble(
  Cont = genefilter_sample(physeq.filt.nct.s.1500.80.unify, filterfun_sample(function(x) x > 0), A=filter*nsamples(physeq.filt.nct.s.1500.80.unify)),
  OTU = names(genefilter_sample(physeq.filt.nct.s.1500.80.unify, filterfun_sample(function(x) x > 0), A=filter*nsamples(physeq.filt.nct.s.1500.80.unify)))
) %>% 
  arrange (-Cont)
cont.count <- filtered_taxa1 %>% count("Cont") 
cont.count %>% 
  kable (caption=paste("Count of contaminants (filter", filter*100, " %)"), booktabs =TRUE) %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) 
```

The list of contaminants determined by high prevalence in NCTs (threhold 0.3)

```{r}
df.fam.melt <- psmelt(physeq.filt.nct.s.1500.80.unify)
df.fam.melt %>% select(OTU, species) %>% 
  unique %>% 
  right_join(filtered_taxa1) %>% 
  filter(Cont == TRUE) %>% 
  select(-Cont) %>% 
  kable (caption=paste("Contaminants (filter", filter*100, " %)"), booktabs =TRUE) %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% 
  scroll_box(width = "70%", height = "600px")
```

We now remove those taxa which have high prevalence in NTC samples out of true samples

```{r}
p.true.RmLowAbun.RmHighPrev.man <- prune_taxa(setdiff(taxa_names(p.true.RmLowAbun.man),filtered_taxa1[filtered_taxa1$Cont==TRUE,] %>% pull("OTU")), p.true.RmLowAbun.man)
```

Now we have *`r nrow(sample_data(p.true.RmLowAbun.RmHighPrev.man))`* true samples and *`r nrow(otu_table(p.true.RmLowAbun.RmHighPrev.man))`* taxa. In NCTs, we have *`r nrow(sample_data(physeq.filt.nct.s.1500.80.unify))`* control samples and *`r nrow(otu_table(physeq.filt.nct.s.1500.80.unify))`* taxa.

The Venn diagram of taxa between true samples and NCTs is now

```{r}
PlotVenn2Sets(rownames(otu_table(p.true.RmLowAbun.RmHighPrev.man)), rownames(otu_table(physeq.filt.nct.s.1500.80.unify)), "True samples", "NCT sample")
```

-- Now Nejman via written code, not function --

```{r}
p.comb.filt.s.1500.80.r <- merge_phyloseq(p.true.RmLowAbun.RmHighPrev.man,physeq.filt.nct.s.1500.80.unify) # merge the phyloseq objects containing true samples (p.true.filt.s.1500.80) and nct samples (physeq.filt.nct.s.1500.80.unify)
p.comb.filt.s.1500.80.r <- tax_glom(p.comb.filt.s.1500.80.r,taxrank = "species", NArm = TRUE) # agglomerate taxa to species level. This combines different OTUs {hypothecially: 450 / 453} assigned to the same species into just 1 OTU {this progess accounts for }
```
#### Loading Batch metadata
```{r introduce Batch metadata}
df <- read_excel("../raw_data/Metadata.xlsx", 
    col_types = c("numeric", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "numeric", "text", "numeric", 
        "numeric", "text", "text")) #load metadata excel-sheet containing DNA extraction batch information & PCR batch infromation
df <- as.data.frame(df)

sd.new <- meta(p.comb.filt.s.1500.80.r) %>% arrange(uniqueID)
df <- df %>% arrange(uniqueID)
df <- df[df$uniqueID %in% sd.new$uniqueID,]

sd.new$DNAex_round <- df$DNAex_round
sd.new$PCR_round <- df$PCR_round

rownames(sd.new) <- sd.new$id

# the above code arranges sampledata of merged phyloseq-object by uniqueID & the metadata excel-sheet by uniqueID. This guarantees the ordering of the rows for both objects is the same. It follows that we can now create the columns "DNAex_round" & "PCR_round" in sd.new using our df. As sample_data from phyloseq-objects have rownames equivalent to the names of their samples we define the rownames of sd.new using the $id colums

p.comb.filt.s.1500.80.r <- merge_phyloseq(otu_table(p.comb.filt.s.1500.80.r), tax_table(p.comb.filt.s.1500.80.r), sample_data(sd.new)) #we merge the otu_table and tax_table of our old phyloseq-object and the newly created sample_data - now containing the DNA extraction batch information and PCR batch information
sample_data(p.comb.filt.s.1500.80.r)$is.neg <- sample_data(p.comb.filt.s.1500.80.r)$TRUE_control == "control"#first identify your negative controls -> decontam!
```

```{r}
#p.comb.filt.s.1500.80.true <- subset_samples(p.comb.filt.s.1500.80.r,TRUE_control=="TRUE") #only true samples
#p.comb.filt.s.1500.80.true <- prune_taxa(x = p.comb.filt.s.1500.80.true,taxa = taxa_names(p.comb.filt.s.1500.80.true)[-(which(taxa_sums(p.comb.filt.s.1500.80.true)==0))]) #only taxa found in true samples

#p.comb.filt.s.1500.80.nct <- subset_samples(p.comb.filt.s.1500.80.r,TRUE_control == "control") #only NCT samples
#p.comb.filt.s.1500.80.nct <- prune_taxa(x = p.comb.filt.s.1500.80.nct,taxa = taxa_names(p.comb.filt.s.1500.80.nct)[-(which(taxa_sums(p.comb.filt.s.1500.80.nct)==0))]) #only taxa found in NCT samples
```

### Binomial exact test
```{r}
p.comb <- p.comb.filt.s.1500.80.r

IDs <- unique(meta(p.comb)$DNAex_round)
df.condition1.meta <- as.data.frame(matrix(0,nrow = 8,ncol = length(IDs)),row.names = c("n_true_samples","n_nct_samples","n_paraffin_tumor_nct_samples","n_paraffin_normal_nct_samples","n_paraffin_nct_samples","n_buffer_nct_samples","n_pcr_nct_samples","batch_name"))
vec0 <- rep(0,length(IDs))
for (i in seq(1,length(IDs))) {
  vec0[i] <- paste0("DNAex","b",i)
}
colnames(df.condition1.meta) <- vec0
for (i in seq(1,length(IDs))) {
  b <- subset_samples(p.comb, DNAex_round == IDs[i])
  b <- prune_taxa(taxa = (taxa_names(b)[-which(taxa_sums(b) == 0)]),x = b) #this removes taxa with taxa_sums == 0 from the overall batch
  check1 <- length(which(meta(b)$TRUE_control == "TRUE")) #checks whether the batch contains true_samples
  check2 <- length(which(meta(b)$TRUE_control == "control")) #checks whether the batch contains nct_sample
  check3 <- length(which(meta(b)$sample_side == "buffer")) #checks whether the batch contains buffer_nct_samples
  check4 <- length(which(meta(b)$sample_side == "PCR_ctrl_H2O")) #checks whether the batch contains pcrH2O_nct_samples
  check5 <- length(which(meta(b)$sample_side == "paraffin_normal")) #checks whether the batch contains paraffinnormal_nct_samples
  check6 <- length(which(meta(b)$sample_side == "paraffin_tumor")) #checks whether the batch contains paraffintumor_nct_samples
  df.condition1.meta[1,i] <- check1
  df.condition1.meta[2,i] <- check2
  df.condition1.meta[3,i] <- check6
  df.condition1.meta[4,i] <- check5
  df.condition1.meta[5,i] <- (check5+check6)
  df.condition1.meta[6,i] <- check3
  df.condition1.meta[7,i] <- check4
  df.condition1.meta[8,i] <- IDs[i]
  if (check1 == 0) {
    print(paste0("check1 :",IDs[i]))
    next}
  if (check2 == 0) {
    print(paste0("check2 :",IDs[i]))
    next}
  if (check3 != 0) {
    #perform test
    b.true <- subset_samples(b,TRUE_control == "TRUE") #subset of true samples in batch
    b.condition <- subset_samples(b,sample_side == "buffer") #subset of batch_nct_samples in batch
    b.sub <- merge_phyloseq(b.true,b.condition) # we take the subsets we want, merge them to allow for thorough removal of bacteria
    if (length(which(taxa_sums(b.sub) == 0)) != 0) {
    b.sub <- prune_taxa(taxa = (taxa_names(b.sub)[-which(taxa_sums(b.sub) == 0)]),x = b.sub) #this removes taxa with taxa_sums == 0 from our batch subset
    }
    b.true <- subset_samples(b.sub,TRUE_control == "TRUE") #create new subset of true samples in batch
    b.condition <- subset_samples(b.sub,sample_side == "buffer") #create new subset of batch_nct_samples in batch  
    # Im honest: this is a bit redundant
    df.filt <- data.frame(x = prevalence(b.true, detection = 0, count = TRUE),
                        n = nsamples(b.true),
                        p = prevalence(b.condition, detection = 0, count = FALSE)) #df stores x,n,p for binom.test
    vec <- seq(1,nrow(df.filt))
    for (j in vec) {
    vec[j] <- binom.test(x = df.filt[j,"x"], n = df.filt[j,"n"], p = df.filt[j,"p"],alternative = "greater")$p.value
    }
    y <- paste0("df.filt1-buffer-",IDs[i])
    assign(y,df.filt)
    df.p <- data.frame(id = rownames(df.filt),
                       pvalue = vec)
    x <- paste0("df.p1-buffer-",IDs[i])
    assign(x,df.p)
  }
  if (check4 != 0) {
    #for DNAex this is mostly irrelevant as PCR_ctrl_H2O has DNAex_round: NA
    #perform test
    b.true <- subset_samples(b,TRUE_control == "TRUE") #subset of true samples in batch
    b.condition <- subset_samples(b,sample_side == "PCR_ctrl_H2O") #subset of batch_nct_samples in batch
    b.sub <- merge_phyloseq(b.true,b.condition) # we take the subsets we want, merge them to allow for thorough removal of bacteria
    if (length(which(taxa_sums(b.sub) == 0)) != 0) {
    b.sub <- prune_taxa(taxa = (taxa_names(b.sub)[-which(taxa_sums(b.sub) == 0)]),x = b.sub) #this removes taxa with taxa_sums == 0 from our batch subset
    }
    b.true <- subset_samples(b.sub,TRUE_control == "TRUE") #create new subset of true samples in batch
    b.condition <- subset_samples(b.sub,sample_side == "PCR_ctrl_H2O") #create new subset of batch_nct_samples in batch  
    # Im honest: this is a bit redundant
    df.filt <- data.frame(x = prevalence(b.true, detection = 0, count = TRUE),
                        n = nsamples(b.true),
                        p = prevalence(b.condition, detection = 0, count = FALSE)) #df stores x,n,p for binom.test
    vec <- seq(1,nrow(df.filt))
    for (j in vec) {
    vec[j] <- binom.test(x = df.filt[j,"x"], n = df.filt[j,"n"], p = df.filt[j,"p"],alternative = "greater")$p.value
    }
    y <- paste0("df.filt1-PCR_ctrl_H2O-",IDs[i])
    assign(y,df.filt)
    df.p <- data.frame(id = rownames(df.filt),
                       pvalue = vec)
    x <- paste0("df.p1-PCR_ctrl_H2O-",IDs[i])
    assign(x,df.p)
  }
  n_paraffin <- (check5+check6)
  if (n_paraffin != 0) {
    #perform test
    b.true <- subset_samples(b,TRUE_control == "TRUE") #subset of true samples in batch
    b.condition <- subset_samples(b,sample_side == "paraffin_tumor" |  sample_side == "paraffin_normal") #subset of batch_nct_samples in batch
    b.sub <- merge_phyloseq(b.true,b.condition) # we take the subsets we want, merge them to allow for thorough removal of bacteria
    if (length(which(taxa_sums(b.sub) == 0)) != 0) {
    b.sub <- prune_taxa(taxa = (taxa_names(b.sub)[-which(taxa_sums(b.sub) == 0)]),x = b.sub) #this removes taxa with taxa_sums == 0 from our batch subset
    }
    b.true <- subset_samples(b.sub,TRUE_control == "TRUE") #create new subset of true samples in batch
    b.condition <- subset_samples(b.sub,sample_side == "paraffin_tumor" |  sample_side == "paraffin_normal") #create new subset of batch_nct_samples in batch  
    # Im honest: this is a bit redundant
    df.filt <- data.frame(x = prevalence(b.true, detection = 0, count = TRUE),
                        n = nsamples(b.true),
                        p = prevalence(b.condition, detection = 0, count = FALSE)) #df stores x,n,p for binom.test
    vec <- seq(1,nrow(df.filt))
    for (j in vec) {
    vec[j] <- binom.test(x = df.filt[j,"x"], n = df.filt[j,"n"], p = df.filt[j,"p"],alternative = "greater")$p.value
    }
    y <- paste0("df.filt1-paraffin-",IDs[i])
    assign(y,df.filt)
    df.p <- data.frame(id = rownames(df.filt),
                       pvalue = vec)
    x <- paste0("df.p1-paraffin-",IDs[i])
    assign(x,df.p)
  }
}

IDs <- unique(meta(p.comb)$PCR_round)
df.condition2.meta <- as.data.frame(matrix(0,nrow = 8,ncol = length(IDs)),row.names = c("n_true_samples","n_nct_samples","n_paraffin_tumor_nct_samples","n_paraffin_normal_nct_samples","n_paraffin_nct_samples","n_buffer_nct_samples","n_pcr_nct_samples","batch_name"))
vec0 <- rep(0,length(IDs))
for (i in seq(1,length(IDs))) {
  vec0[i] <- paste0("PCR","b",i)
}
colnames(df.condition2.meta) <- vec0
for (i in seq(1,length(IDs))) {
  b <- subset_samples(p.comb, PCR_round == IDs[i])
  b <- prune_taxa(taxa = (taxa_names(b)[-which(taxa_sums(b) == 0)]),x = b) #this removes taxa with taxa_sums == 0 from the overall batch
  check1 <- length(which(meta(b)$TRUE_control == "TRUE")) #checks whether the batch contains true_samples
  check2 <- length(which(meta(b)$TRUE_control == "control")) #checks whether the batch contains nct_sample
  check3 <- length(which(meta(b)$sample_side == "buffer")) #checks whether the batch contains buffer_nct_samples
  check4 <- length(which(meta(b)$sample_side == "PCR_ctrl_H2O")) #checks whether the batch contains pcrH2O_nct_samples
  check5 <- length(which(meta(b)$sample_side == "paraffin_normal")) #checks whether the batch contains paraffinnormal_nct_samples
  check6 <- length(which(meta(b)$sample_side == "paraffin_tumor")) #checks whether the batch contains paraffintumor_nct_samples
  df.condition2.meta[1,i] <- check1
  df.condition2.meta[2,i] <- check2
  df.condition2.meta[3,i] <- check6
  df.condition2.meta[4,i] <- check5
  df.condition2.meta[5,i] <- (check5+check6)
  df.condition2.meta[6,i] <- check3
  df.condition2.meta[7,i] <- check4
  df.condition2.meta[8,i] <- IDs[i]
  if (check1 == 0) {
    print(paste0("check1 :",IDs[i]))
    next}
  if (check2 == 0) {
    print(paste0("check2 :",IDs[i]))
    next}
  if (check3 != 0) {
    #perform test
    b.true <- subset_samples(b,TRUE_control == "TRUE") #subset of true samples in batch
    b.condition <- subset_samples(b,sample_side == "buffer") #subset of batch_nct_samples in batch
    b.sub <- merge_phyloseq(b.true,b.condition) # we take the subsets we want, merge them to allow for thorough removal of bacteria
    if (length(which(taxa_sums(b.sub) == 0)) != 0) {
    b.sub <- prune_taxa(taxa = (taxa_names(b.sub)[-which(taxa_sums(b.sub) == 0)]),x = b.sub) #this removes taxa with taxa_sums == 0 from our batch subset
    }
    b.true <- subset_samples(b.sub,TRUE_control == "TRUE") #create new subset of true samples in batch
    b.condition <- subset_samples(b.sub,sample_side == "buffer") #create new subset of batch_nct_samples in batch  
    # Im honest: this is a bit redundant
    df.filt <- data.frame(x = prevalence(b.true, detection = 0, count = TRUE),
                        n = nsamples(b.true),
                        p = prevalence(b.condition, detection = 0, count = FALSE)) #df stores x,n,p for binom.test
    vec <- seq(1,nrow(df.filt))
    for (j in vec) {
    vec[j] <- binom.test(x = df.filt[j,"x"], n = df.filt[j,"n"], p = df.filt[j,"p"],alternative = "greater")$p.value
    }
    y <- paste0("df.filt2-buffer-",IDs[i])
    assign(y,df.filt)
    df.p <- data.frame(id = rownames(df.filt),
                       pvalue = vec)
    x <- paste0("df.p2-buffer-",IDs[i])
    assign(x,df.p)
  }
  if (check4 != 0) {
    #for DNAex this is mostly irrelevant as PCR_ctrl_H2O has DNAex_round: NA
    #perform test
    b.true <- subset_samples(b,TRUE_control == "TRUE") #subset of true samples in batch
    b.condition <- subset_samples(b,sample_side == "PCR_ctrl_H2O") #subset of batch_nct_samples in batch
    b.sub <- merge_phyloseq(b.true,b.condition) # we take the subsets we want, merge them to allow for thorough removal of bacteria
    if (length(which(taxa_sums(b.sub) == 0)) != 0) {
    b.sub <- prune_taxa(taxa = (taxa_names(b.sub)[-which(taxa_sums(b.sub) == 0)]),x = b.sub) #this removes taxa with taxa_sums == 0 from our batch subset
    }
    b.true <- subset_samples(b.sub,TRUE_control == "TRUE") #create new subset of true samples in batch
    b.condition <- subset_samples(b.sub,sample_side == "PCR_ctrl_H2O") #create new subset of batch_nct_samples in batch  
    # Im honest: this is a bit redundant
    df.filt <- data.frame(x = prevalence(b.true, detection = 0, count = TRUE),
                        n = nsamples(b.true),
                        p = prevalence(b.condition, detection = 0, count = FALSE)) #df stores x,n,p for binom.test
    vec <- seq(1,nrow(df.filt))
    for (j in vec) {
    vec[j] <- binom.test(x = df.filt[j,"x"], n = df.filt[j,"n"], p = df.filt[j,"p"],alternative = "greater")$p.value
    }
    y <- paste0("df.filt2-PCR_ctrl_H2O-",IDs[i])
    assign(y,df.filt)
    df.p <- data.frame(id = rownames(df.filt),
                       pvalue = vec)
    x <- paste0("df.p2-PCR_ctrl_H2O-",IDs[i])
    assign(x,df.p)
  }
  n_paraffin <- (check5+check6)
  if (n_paraffin != 0) {
    #perform test
    b.true <- subset_samples(b,TRUE_control == "TRUE") #subset of true samples in batch
    b.condition <- subset_samples(b,sample_side == "paraffin_tumor" |  sample_side == "paraffin_normal") #subset of batch_nct_samples in batch
    b.sub <- merge_phyloseq(b.true,b.condition) # we take the subsets we want, merge them to allow for thorough removal of bacteria
    if (length(which(taxa_sums(b.sub) == 0)) != 0) {
    b.sub <- prune_taxa(taxa = (taxa_names(b.sub)[-which(taxa_sums(b.sub) == 0)]),x = b.sub) #this removes taxa with taxa_sums == 0 from our batch subset
    }
    b.true <- subset_samples(b.sub,TRUE_control == "TRUE") #create new subset of true samples in batch
    b.condition <- subset_samples(b.sub,sample_side == "paraffin_tumor" |  sample_side == "paraffin_normal") #create new subset of batch_nct_samples in batch  
    # Im honest: this is a bit redundant
    df.filt <- data.frame(x = prevalence(b.true, detection = 0, count = TRUE),
                        n = nsamples(b.true),
                        p = prevalence(b.condition, detection = 0, count = FALSE)) #df stores x,n,p for binom.test
    vec <- seq(1,nrow(df.filt))
    for (j in vec) {
    vec[j] <- binom.test(x = df.filt[j,"x"], n = df.filt[j,"n"], p = df.filt[j,"p"],alternative = "greater")$p.value
    }
    y <- paste0("df.filt2-paraffin-",IDs[i])
    assign(y,df.filt)
    df.p <- data.frame(id = rownames(df.filt),
                       pvalue = vec)
    x <- paste0("df.p2-paraffin-",IDs[i])
    assign(x,df.p)
  }
}

bad_taxa <- list()
#bad_taxa <- list.append(bad_taxa,`df.p1-buffer-1`[which(x = `df.p1-buffer-1`$pvalue >= 0.05),1])
# bad_taxa <- list.append(bad_taxa,`df.p1-buffer-2`[which(x = `df.p1-buffer-2`$pvalue >= 0.05),1])
# bad_taxa <- list.append(bad_taxa,`df.p1-buffer-3`[which(x = `df.p1-buffer-3`$pvalue >= 0.05),1])
# bad_taxa <- list.append(bad_taxa,`df.p1-buffer-4`[which(x = `df.p1-buffer-4`$pvalue >= 0.05),1])
# bad_taxa <- list.append(bad_taxa,`df.p1-buffer-5`[which(x = `df.p1-buffer-5`$pvalue >= 0.05),1])
#bad_taxa <- list.append(bad_taxa,`df.p1-buffer-6`[which(x = `df.p1-buffer-6`$pvalue >= 0.05),1])
bad_taxa <- list.append(bad_taxa,`df.p1-paraffin-1`[which(x = `df.p1-paraffin-1`$pvalue >= 0.05),1])
bad_taxa <- list.append(bad_taxa,`df.p1-paraffin-2`[which(x = `df.p1-paraffin-2`$pvalue >= 0.05),1])
bad_taxa <- list.append(bad_taxa,`df.p1-paraffin-3`[which(x = `df.p1-paraffin-3`$pvalue >= 0.05),1])
bad_taxa <- list.append(bad_taxa,`df.p1-paraffin-4`[which(x = `df.p1-paraffin-4`$pvalue >= 0.05),1])
bad_taxa <- list.append(bad_taxa,`df.p1-paraffin-5`[which(x = `df.p1-paraffin-5`$pvalue >= 0.05),1])
#bad_taxa <- list.append(bad_taxa,`df.p1-paraffin-6`[which(x = `df.p1-paraffin-6`$pvalue >= 0.05),1])

#bad_taxa <- list.append(bad_taxa,`df.p2-buffer-8`[which(x = `df.p2-buffer-8`$pvalue >= 0.05),1])
#bad_taxa <- list.append(bad_taxa,`df.p2-PCR_ctrl_H2O-1`[which(x = `df.p2-PCR_ctrl_H2O-1`$pvalue >= 0.05),1])
bad_taxa <- list.append(bad_taxa,`df.p2-PCR_ctrl_H2O-2`[which(x = `df.p2-PCR_ctrl_H2O-2`$pvalue >= 0.05),1])
#bad_taxa <- list.append(bad_taxa,`df.p2-paraffin-8`[which(x = `df.p2-paraffin-8`$pvalue >= 0.05),1])

# a <- union(bad_taxa[[1]],bad_taxa[[2]])
# b <- union(bad_taxa[[3]],bad_taxa[[4]])
# c <- union(bad_taxa[[5]],bad_taxa[[6]])
# d <- union(bad_taxa[[7]],bad_taxa[[8]])
# e <- union(bad_taxa[[9]],bad_taxa[[10]])
# f <- union(bad_taxa[[11]],bad_taxa[[12]])
# g <- union(bad_taxa[[13]],bad_taxa[[14]])
# h <- union(bad_taxa[[15]],bad_taxa[[16]])
# 
# a <- union(a,b)
# c <- union(c,d)
# e <- union(e,f)
# g <- union(g,h)
# 
# a <- union(a,c)
# e <- union(e,g)

bad_taxa <- unique(unlist(bad_taxa)) #all OTUs representing species we have to remove from p.true.RmLowAbun.RmHighPrev.man
length(bad_taxa)

# using the following commentended code is only correct if the OTUs from bad_taxa have the same reference as our p.true.RmLowAbun.RmHighPrev.man phyloseq object 
#p.true.RmLowAbun.RmHLPrev.man1 <- subset_taxa(p.true.RmLowAbun.RmHighPrev.man,!(taxa_names(p.true.RmLowAbun.RmHighPrev.man) %in% bad_taxa))

p.comb <- subset_taxa(p.comb, !(taxa_names(p.comb) %in% bad_taxa))
p.true.RmLowAbun.RmHLPrev.man <- subset_samples(physeq = p.comb, TRUE_control == "TRUE")
p.true.RmLowAbun.RmHLPrev.man <- prune_taxa(x = p.true.RmLowAbun.RmHLPrev.man,taxa = taxa_names(p.true.RmLowAbun.RmHLPrev.man)[-(which(taxa_sums(p.true.RmLowAbun.RmHLPrev.man)==0))])
```

The overlap is now

```{r}
PlotVenn2Sets(p.true.RmLowAbun.RmHLPrev.man %>% otu_table() %>% rownames(),
              physeq.filt.nct.s.1500.80.unify %>% otu_table() %>% rownames(),
              "True samples", "NCT samples")
```

Further, we have a look at those 49 taxas in NCTs but not contaminants

```{r}
physeq.nct.NotCont.man <- prune_taxa(intersect(physeq.filt.nct.s.1500.80.unify %>% otu_table() %>% rownames(), 
                                           p.true.RmLowAbun.RmHLPrev.man %>% otu_table() %>% rownames()),
                                 physeq.filt.nct.s.1500.80.unify)
Inspect_taxa_Species(physeq.nct.NotCont.man, "In NCTs but not contaminants")
```

### Inter-Batch effect correction

We apply ConQuR to correct batch effect across batches.

First, we apply the batch correction on DNAex_round

```{r}
#p.true.RmLowAbun.RmHLPrev.batchDNA <- ConQuR_applyCP(phyloseq = p.true.RmLowAbun.RmHLPrev,meta_data = metadata,batch_type = "DNAex",batch_ref = 1)
batch_type = "DNA"
otu <- as(otu_table(p.true.RmLowAbun.RmHLPrev.man), "matrix")
if(taxa_are_rows(p.true.RmLowAbun.RmHLPrev.man)){otu <- t(otu)}
 if(grepl("dna", tolower(batch_type), fixed = TRUE)){
  sub.meta.data <- sd.new[sd.new$TRUE_control == "TRUE",]
  sub.meta.data <- sub.meta.data %>% mutate(sample_sidev2 = as.factor(sample_side),
                                               DNAex_roundv2 = as.factor(DNAex_round))
  batchid = sub.meta.data[,c("DNAex_roundv2")]
  covar = batchid
  batch_refer = "2"
}
 if(grepl("pcr", tolower(batch_type), fixed = TRUE)){
   sub.meta.data <- sd.new[sd.new$TRUE_control == "TRUE",]
   sub.meta.data <- sub.meta.data %>% mutate(sample_sidev2 = as.factor(sample_side),
                                               PCR_roundv2 = as.factor(PCR_round))
   batchid = sub.meta.data[,c("PCR_roundv2")]
   covar = batchid
   batch_refer = "2"
}
options(warn = -2)
taxa_corrected <- ConQuR(tax_tab = otu,batchid = batchid,covariates = covar,batch_ref = batch_refer)
tmp_phy <- p.true.RmLowAbun.RmHLPrev.man
otu_table(tmp_phy) <- otu_table(taxa_corrected %>% t(),taxa_are_rows = TRUE)
Plot_PCoA(TAX = otu,factor = batchid,main="Pre-Correction, Bray Curtis")
Plot_PCoA(TAX = taxa_corrected,factor = batchid ,main = "Post-Correction, Bray Curtis")
p.true.RmLowAbun.RmHLPrev.batchDNA.man <- tmp_phy
rm(batch_type)
rm(otu)
rm(sub.meta.data)
rm(batchid)
rm(covar)
rm(batch_refer)
rm(tmp_phy)
```

```{r}
#p.true.RmLowAbun.RmHLPrev.batchPCR <- ConQuR_applyCP(phyloseq = p.true.RmLowAbun.RmHLPrev.batchDNA,meta_data = metadata,batch_type = "PCR",batch_ref = 1)

batch_type = "PCR"
otu <- as(otu_table(p.true.RmLowAbun.RmHLPrev.batchDNA.man), "matrix")
if(taxa_are_rows(p.true.RmLowAbun.RmHLPrev.batchDNA.man)){otu <- t(otu)}
if(grepl("dna", tolower(batch_type), fixed = TRUE)){
  sub.meta.data <- sd.new[sd.new$TRUE_control == "TRUE",]
  sub.meta.data <- sub.meta.data %>% mutate(sample_sidev2 = as.factor(sample_side),
                                              DNAex_roundv2 = as.factor(DNAex_round))
  batchid = sub.meta.data[,c("DNAex_roundv2")]
  covar = sub.meta.data[,c("sample_sidev2")]
  batch_refer = "1"
}
if(grepl("pcr", tolower(batch_type), fixed = TRUE)){
  sub.meta.data <- sd.new[sd.new$TRUE_control == "TRUE",]
  sub.meta.data <- sub.meta.data %>% mutate(sample_sidev2 = as.factor(sample_side),
                                              PCR_roundv2 = as.factor(PCR_round))
  batchid = sub.meta.data[,c("PCR_roundv2")]
  covar = sub.meta.data[,c("sample_sidev2")]
  batch_refer = "2"
}
options(warn = -2)
taxa_corrected <- ConQuR(tax_tab = otu,batchid = batchid,covariates = covar,batch_ref = batch_refer)
tmp_phy <- p.true.RmLowAbun.RmHLPrev.batchDNA.man
otu_table(tmp_phy) <- otu_table(taxa_corrected %>% t(),taxa_are_rows = TRUE)
Plot_PCoA(TAX = otu,factor = batchid,main="Pre-Correction, Bray Curtis")
Plot_PCoA(TAX = taxa_corrected,factor = batchid ,main = "Post-Correction, Bray Curtis")
p.true.RmLowAbun.RmHLPrev.batchPCR.man <- tmp_phy
rm(batch_type)
rm(otu)
rm(sub.meta.data)
rm(batchid)
rm(covar)
rm(batch_refer)
rm(tmp_phy)
```
--

```{r}
p.true.RmLowAbun.RmHLPrev.batch.man <- p.true.RmLowAbun.RmHLPrev.batchPCR.man
```


### Normalization by Rarefaction

First of all, we have a look at bacterial reads of each sample

```{r}
Inspect_SequencingDepth(p.true.RmLowAbun.RmHLPrev.batch.man)
```


Rarefying curve

```{r}
tab <- p.true.RmLowAbun.RmHLPrev.batch.man %>% otu_table() 
class(tab) <- "matrix"
rarecurve(tab %>% t(), step=50, cex=0.5)
```




```{r}
#p.comb.filt.s.1500.80.true <- subset_samples(p.comb.filt.s.1500.80.r,TRUE_control=="TRUE") #only true samples
#p.comb.filt.s.1500.80.true <- prune_taxa(x = p.comb.filt.s.1500.80.true,taxa = taxa_names(p.comb.filt.s.1500.80.true)[-(which(taxa_sums(p.comb.filt.s.1500.80.true)==0))]) #only taxa found in true samples
```
### Prev-Prev plot
```{r}
# # Make phyloseq object of presence-absence in negative controls and true samples
# ps.pa <- transform_sample_counts(p.comb.filt.s.1500.80, function(abund) 1*(abund>0)) #based on original data
# ps.pa.neg <- prune_samples(sample_data(ps.pa)$TRUE_control == "control", ps.pa)
# ps.pa.pos <- prune_samples(sample_data(ps.pa)$TRUE_control == "TRUE", ps.pa)
# # Make data.frame of prevalence in positive and negative samples
# 
# NoContaminant.nej <- taxa_names(p.comb.filt.s.1500.80) %in% taxa_names(p.true.RmLowAbun.RmHLPrev.batchPCR.man)
# NoContaminant.nej <- as.character(NoContaminant.nej)
# df.pa <- data.frame(pa.pos=taxa_sums(ps.pa.pos), pa.neg=taxa_sums(ps.pa.neg),
#                       NoContaminant=NoContaminant.nej)
# ggplot(data=df.pa, aes(x=pa.neg, y=pa.pos, color=NoContaminant)) + geom_point() +
#   xlab("Prevalence (Negative Controls)") + ylab("Prevalence (True Samples)")
# 
# #Observe taxa with Prevalence in True Sample of 0 are still contained in the data -> not tested by our method
```

### Decontam p-Histogram
```{r}
# noncontamdf.prev <- isNotContaminant(p.comb.filt.s.1500.80.r, method = "prevalence", neg ="is.neg",threshold = 0.5,detailed = TRUE)
# #threshold of 0.5 classifies all taxa present in more negative controls then tisse samples as contaminant
# #intersting would be the comparison of this classification and our NEJMAN derived classification method
# hist(noncontamdf.prev$p, 100, ylim = c(0,300), xlim = c(0,1))
```
